The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2017-08-22 JStatSoft_forestinventory_master-concordance.tex
update Andreas 22.08.2017

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 121 1 1 4 1 1 1 4 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 128:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 224:%
136 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 259:%
1 84 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 344:%
172 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 360:%
1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 156 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 888:%
189}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 889:%
1 92 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 982:%
191 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 993:%
1 13 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1007:%
203 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1016:%
1 20 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1037:%
213 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1047:%
1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 24 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 79 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1289:%
224 20 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 123 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
139 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
1 78 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 341:%
175 13 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Onephase_Estimators and their Application.Rnw:ofs 355:%
1 25 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 381:%
190}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 382:%
1 124 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 911:%
192}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 912:%
1 118 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 179 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1302:%
194 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1313:%
1 13 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1327:%
206 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1336:%
1 21 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1358:%
216 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1374:%
1 20 1 1 4 6 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1629:%
233 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1633:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1638:%
238 11 1}

### 2017-08-22 rnw/Onephase_Estimators and their Application.Rnw
update Andreas 22.08.2017

\section{Threephase Estimators and their Application}
\label{sec:threephase_and_appl}

% ---------------------------------------------------------------------------- %
% ---------------------------------------------------------------------------- %
\subsection{Global Estimators}

% - Merge Mathematical Background and Application
% - Quite Likely, we cannot go through all estimators ...
%   Idea: --> maybe only for non-exhaustive case, and then demonstrate with one example how the exhaustive can be applied in general
%
% 
% NOTES:
% - mention that the exhaustive 3phase was originally called "2phase with partially exhaustive info"
% - clarify terminology "full" and "reduced"
%

% ------------------------------------- %
\subsubsection{Mathematical Background}

The the vector of regression coefficients for the \textbf{\textit{reduced model}} using the reduced set of explanatory variables $\pmb{Z}^{(0)}(x)$ is derived by solving the normal equation

\begin{eqnarray}\label{eq:normequ_redmod}
\hat{\pmb{\alpha}}_{s_2} &=& \Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}^{(0)}(x)
\end{eqnarray}

and likewise the vector of regression coefficients for the \textbf{\textit{full model}} using the full set of explanatory variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))$:

\begin{eqnarray}\label{eq:normequ_redmod}
\hat{\pmb{\beta}}_{{s_2}}&=&\Big(\frac{1}{n_2}\sum_{x\in{s}_2}\pmb{Z}(x)\pmb{Z}^t(x)
\Big)^{-1}\frac{1}{n_2}\sum_{x\in{s}_2}Y(x)\pmb{Z}(x)
\end{eqnarray}

The variance-covariance matrices of the regression coefficients are thus defined as

\begin{subequations}\label{eq:covar3p}
\begin{align}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x)\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x)\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}^{(0)}(x)\pmb{Z}^{(0)t}(x) \Big)^{-1} \label{eq:covar_alpha} \\
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}^t(x)\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \label{eq:covar_beta}
\end{align}
\end{subequations}

with the residuals $R_0(x)$ and $R(x)$ derived under the reduced and the full regression model respectively.

Likewise the two-phase estimation (section \ref{sec:twophase_and_appl}), also the three-phase estimation can be applied under \textit{non-exhaustive} and \textit{exhaustive} derivation of the explanatory variables at the $s_0$ phase.

In the \textbf{non-exhaustive case}, we use the \textit{estimated} means of $\pmb{Z}^{(0)}(x)$ over both the $s_0$ and $s_1$ phase, as well as the \textit{estimated} mean of $\pmb{Z}(x)$ over both the $s_1$ phase (equation \ref{meanvalues3pnex}). If the explanatory variables used in the reduced model are exhaustively derived (\textbf{exhaustive case}), the estimated mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ is simply replaced by the exact mean $\bar{\pmb{Z}}^{(0)}_0$. Note that in case of applied boundary adjustment (section \ref{sec:twophase_and_appl}), the simple mean is again replaced by the weighted mean analogous to equation \ref{eq:wmeanZ}.

\begin{equation}\label{meanvalues3pnex}
\hat{\bar{\pmb{Z}}}^{(0)}_0=\frac{1}{n_0}\sum_{x\in{s_0}} \pmb{Z}^{(0)}(x), \quad \hat{\bar{\pmb{Z}}}^{(0)}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}^{(0)}(x) ,
\quad \hat{\bar{\pmb{Z}}}_1=\frac{1}{n_1}\sum_{x\in{s}_1}\pmb{Z}(x)
\end{equation}

Under the internal model prerequisite which insures that the zero mean residual property over $F$, \textbf{point estimates} are then calculated as

\begin{subequations}\label{reg3p}
\begin{align}
\hat{Y}_{reg3p,ex}&=(\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_ex} \\
\hat{Y}_{reg3p,nex}&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^t\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{2} \label{eq:reg3p_nex}
\end{align}
\end{subequations}

Equations \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} nicely demonstrate the philosophy of the three-phase estimation approach. Without the first term, the equations exactly equal the one of two-phase estimator where the estimated means in $\hat{\bar{\pmb{Z}}}_1$ are based on the $s_1$ sample and are transformed into predictions $\hat{Y}(x)$ of the local density $Y(x)$. The difference however is that we can now derive a more precise mean $\hat{\bar{\pmb{Z}}}^{(0)}_0$ (or even the exact mean $\bar{\pmb{Z}}^{(0)}_0$) over $s_0$ for a subset $\pmb{Z}^{(0)}(x) \in \pmb{Z}(x)$. The first term of equation \ref{eq:reg3p_nex} and \ref{eq:reg3p_ex} can thus be interpreted as a correction term that transforms the difference to the more precise / exact mean of $\pmb{Z}^{(0)}(x)$ into a residual of $Y(x)$ and adds it to the second term, thus making the point estimate more precise.

The \textbf{design-based variance} is then calculated as

\begin{subequations}\label{eq:dbvar_reg3p}
\begin{align}
\hat{\var}(\hat{Y}_{reg3p,ex})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)t}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_1^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_ex} \\
\hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
+\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}_1 \label{eq:dbvar_reg3p_nex}
\end{align}
\end{subequations}

with the variance-covariance matrix $\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(1)}_0}$:

\begin{equation}\label{estvarcovaux3p}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}=
\frac{1}{n_{0}(n_{0}-1)}\sum_{x\in{s_{0}}}
(\pmb{Z}^{(0)}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}^{(0)}_{0})^t
\end{equation}

We see that equation \ref{eq:dbvar_reg3p_ex} (exhaustive case) and \ref{eq:dbvar_reg3p_nex} (non-exhaustive case) differ by the term $\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2$ which is exactly the variance of the prediction $\hat{Y}_{0}(x)$ of the reduced model in the $s_0$ phase. This term can be made very small by chosing a large sample size $n_0$ for $s_0$ \citep{mandallaz2013techb}.

The package also gives the \textbf{external variance} defined as

\begin{subequations}\label{eq:extvar_reg3p}
\begin{align}
\hat{\var}(\hat{Y}_{reg3p,ex})&=\frac{1}{n_1}\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}_{0}^2(x))
+ (1-\frac{n_2}{n_1})\frac{1}{n^2_2}\sum_{x\in{s_2}}\hat{R}^2(x) \label{eq:extvar_reg3p_ex} \\
\hat{\var}_{ext}(\hat{Y}_{reg3p,nex})&=\frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_0(x)-\hat{\bar{Y}}_0)^2}{n_0-1} 
+\frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_0(x)-\hat{\bar{R_0}})^2 \nonumber\\
&+(1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
\end{align}
\end{subequations}

with ${R}(x)$ being the full model residuals calculated as $Y(x)-\pmb{Z}^t(x)\hat{\beta}_{s_2}$ and ${R}_{0}(x)$ being the reduced model residuals calculated as $Y(x)-\pmb{Z}^{(0)t}\hat{\alpha}_{s_2}$.

% \begin{subequations}\label{eq:var_reg3p_nex}
% \begin{align}
% \hat{\var}(\hat{Y}_{reg3p,nex})& =\hat{\pmb{\alpha}}_2^t\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_0}\hat{\pmb{\alpha}}_2
% +\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)t}_0
% \hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{2}}\hat{\bar{\pmb{Z}}}^{(0)}_0 + (1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}^{t}_1\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_2}\hat{\bar{\pmb{Z}}}^{t}_1 \label{eq:dbvar_reg3p_nex} \\
% \hat{\var}_{ext}(\hat{Y}_{reg3p,nex})& = \frac{1}{n_0}\frac{\sum_{x\in{s_0}}(\hat{Y}_1(x)-\hat{\bar{Y}}_1)^2}{n_0-1} \nonumber\\
% &+ \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s}_2}(\hat{R}_1(x)-\hat{\bar{R_1}})^2+
% (1-\frac{n_2}{n_1})\frac{1}{n_2(n_2-1)}\sum_{x\in{s}_2}(\hat{R}(x)-\hat{\bar{R}})^2 \label{eq:extvar_reg3p_nex}
% \end{align}
% \end{subequations}

% ------------------------------------- %
\subsubsection{Application}

%--------------------------------------------------------------------------------------------------%
% ################################################################################################ %
%--------------------------------------------------------------------------------------------------%

% ---------------------------------------------------------------------------- %
\subsection{Small Area Estimators}

% ------------------------------------- %
\subsubsection{Mathematical Background}

%------------------------------------------------------------- %
\subsubsection{Application}

>>>>

\section{One-phase Estimators and their Application}
\label{sec:onephase_and_appl}

% ---------------------------------------------------------------------------- %
  \subsection{One-phase Estimators}

\subsubsection{Mathematical Background}

\subsubsection{Application}
